THESIS PROJECT PROPOSAL

### 2019-11-15 project_ashish/paper.tex
proposal, objetivo, hypthesis

\begin{figure}[H]
  \centering
   \includegraphics[scale=0.5]{./cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels.}
   \label{figure 3}
\end{figure}
\clearpage
}
% \newpage

@@

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{./cms.png}
	\caption{Leading order Feynman diagrams for the associated production of a single top quark and a Higgs boson in the t channel, where the Higgs boson couples either to the top quark (left) or the W boson (right).}
	\label{figure 4}
\end{figure}

@@

\afterpage{

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.36]{./cms12.png}
	\caption{Transverse view of the CMS detector showing the silicon tracker, superconducting solenoid, electromagnetic calorimeter, hadron calorimeter and muon chambers.}
	\label{figure5}
\end{figure}

 \begin{figure}[H]
 	\centering
 	\includegraphics[scale=0.45]{./lum6.png}
 	\caption{Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points show instant brightness while the line shows brightness accumulated.}
 	\label{figure6}
 \end{figure}

\clearpage
}

\afterpage{

 \begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{./cd7.png}
	\caption{ATLAS-CMS combined measurements of coupling strength modifiers.}
	\label{figureKappas}
\end{figure}

 \begin{figure}[H]
  \centering
   \includegraphics[width=0.5 \columnwidth]{./cms13.png}
  \caption{Scan of -2 $\Delta$ ln(L) versus $\kappa_t$ for the data (black line) and the individual channels (blue, red, and green), compared to Asimov data sets corresponding to the SM expectations (dashed lines).}
   \label{figureKt}
\end{figure}

\clearpage
}

LHC during the first run in 2011 and 2012 reached a peak luminosity of 7.7 $\times$ $10^{33}$ $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 25 $fb^{-1}$ to each ATLAS and CMS.
Figure~\ref{figureKappas} shows a summary of the current measurments of the Higgs coupling parameters with the LHC Run 1 dataset.
All the hadron colliders in the world have so far produced a total integrated luminosity of about 10 $fb^{-1}$, while the LHC will deliver about 300 $fb^{-1}$ by 2024 \cite{collaborations2019report}.
The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles.
In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  placein parallel during one shutdown, at around 2023-25 (Long shutdown 3), with some of the modification anticipated in 2019-2020 (LS2).
This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ by the year 2036 as shown in Figure~\ref{figure6}.

Earlier searches for Higgs boson production production in association with single top quark have been studied to derive constraints on the magnitude and relative sign of Higgs boson couplings to top quarks and vector bosons.
Constaints on $\kappa_t$ can be derived using a likelihood ratio scan of $L(\kappa_t)/L(\hat{\kappa_t})$ where $\hat{\kappa_t}$ is the best fit value of $\kappa_t$ as shown in Figure~\ref{figureKt}.
Multilepton final state and final states with single lepton with a pair of bottom quarks were combined with Higgs decay to two photons to get the final result. The data favour positive value of top quark yukawa coupling for SM Higgs boson and excludes ranges of about [-0.9,-0.5] and [1.0,2.1] times $y^{SM}_t$ at the 95 $\%$ confidence level.

\section{PROPOSAL}

- what we will do:  search for tH events, using which channels.

In this project, it is proposed to study the production channel of the Higgs boson associated 
with a single top quark in the Large Hadron Collider of the CERN laboratory. The measurement 
of this process complements other measurements of the yukawa coupling parameter of the top 
quark and Higgs boson. It is a part of the physics program of CMS international 
collaborations and ATLAS and it has not been observed so far. It is proposed to study 
sensitivity with the current data in the channel where the events are identified with two 
muons of the same sign. The impact of uncertainties will be studied and projections will be 
made to the future phases of the LHC. For these studies the data of the run two (2015-2018) 
of the LHC identifying the signal events with two muons from the same sign. It is also 
expected to estimate projections for the high brightness phase of the LHC (HL-LHC) that will begin 
in 2026.

@@

- motivation, tH production crossection, sign of $k_t$, search for BSM, ..

There is currently a great enigma in physics since we know that ordinary matter comprises 
only $5\%$ of the universe, another 
$27\%$ includes what we call dark matter. Understanding the production of the Higgs boson, as 
well as its decays like the invisible decay modes are an important part of the physics 
program of CERN laboratory experiments to verify the Standard Model and look for new 
particles. Through this project we will investigate the production of the Higgs boson in 
association with a single top quark (tH) in proton-proton collisions with the CMS experiment 
of the LHC. We will study the distribution of events in one of the decay channels and the 
sensitivity that we get with the current data. The main objective is to search for new 
processes such as tH Higgs production mode in the CMS experiment which require to study 
different aspects such as the model of the noises that are made by simulation, techniques of 
adjustments to the data that incorporate the different noises and the signal, and estimates 
of statistical and systematic uncertainties.
Some of the objectives in this project include:
\begin{itemize}
\item study one or more decay channels.
\item verify that simulation of signal events is adequate.
\item study the extraction of signal strength through adjustments to the data.
\item study the impact of statistical and systematic uncertainties.
\end{itemize}

@@

\section{HYPOTHESIS}
- what we expect to find

Production studies in the tH channel have already started with the collaborations CMS and 
ATLAS with publications based on data from the 2011-2012 run and some data from the second 
run of 2015-2016. In these results it was not found signal evidence and upper limits have 
been set for the effective section. With the more sensitive analysis an upper limit is 
calculated for the effective section of a signal composed of tH and ttH that corresponds 3.1 
times that expected in the SM. In this analysis the reconstruction methods do not allow to 
distinguish between tH and t$\bar{t}$H because the effective section of t$\bar{t}$H is much 
larger than that of tH and some events pass the selections designed for tH, the fraction of 
tH events over tH + t$\bar{t}$H is 5. In BSM models the fraction could change up to 50, for 
example with $\kappa_t = -1$. It could help to obtain better limits for the production of tH. 
The data that has been used so far correspond to only 20 of the final data of the run that 
which ends in 2018. The uncertainty for the signal in the current analysis is mostly 
statistical nature, therefore adding the total data in a new analysis the next year, 
sensitivity can be improved with an approximate factor of 2.2. Additionally, in subsequent 
phases of the LHC 2021-2023 and 2026-2038 the number of data fold and then multiply by 10 
respectively. With these additions It is expected that the signal from the Higgs boson can be 
observed on the tH production channel. The current results can be used to project sensitivity 
in the third run and the high brightness phase and thus contribute to the planning of the 
updates of the detector.

>>>>

\begin{figure}[H]
  \centering
   \includegraphics[scale=0.7]{./cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels.}
   \label{figure 3}
\end{figure}
\clearpage
}
% \newpage

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{./cms.png}
	\caption{Leading order Feynman diagrams for the associated production of a single top quark and a Higgs boson in the t channel, where the Higgs boson couples either to the top quark (left) or the W boson (right).}
	\label{figure 4}
\end{figure}

@@

@@

\afterpage{
  \begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\columnwidth]{./cms12.png}
    \caption{Transverse view of the CMS detector showing the silicon tracker, superconducting solenoid, electromagnetic calorimeter, hadron calorimeter and muon chambers.}
    \label{figure5}
  \end{figure}

  \begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\columnwidth]{./lum6.png}
    \caption{Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points show instant brightness while the line shows brightness accumulated.}
    \label{figure6}
  \end{figure}
  \clearpage
}

\afterpage{
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{./cd7.png}
    \caption{ATLAS-CMS combined measurements of coupling strength modifiers.}
    \label{figureKappas}
  \end{figure}

  \begin{figure}[ht]
    \centering
    \includegraphics[width=0.5 \columnwidth]{./cms13.png}
    \caption{Scan of -2 $\Delta$ ln(L) versus $\kappa_t$ for the data (black line) and the individual channels (blue, red, and green), compared to Asimov data sets corresponding to the SM expectations (dashed lines).}
    \label{figureKt}
  \end{figure}
\clearpage
}

LHC during the first run in 2011 and 2012 reached a peak luminosity of 7.7 $\times$ $10^{33}$ $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 25 $fb^{-1}$ to each ATLAS and CMS.
All the hadron colliders in the world have so far produced a total integrated luminosity of about 10 $fb^{-1}$, while the LHC will deliver about 300 $fb^{-1}$ by 2024 \cite{collaborations2019report}.
The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles.
In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  placein parallel during one shutdown, at around 2023-25 (Long shutdown 3), with some of the modification anticipated in 2019-2020 (LS2).
This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ by the year 2036 as shown in Figure~\ref{figure6}.

Figure~\ref{figureKappas} shows a summary of the current measurments of the Higgs coupling parameters with the LHC Run 1 dataset (2011,2012)  using ggF, VBF, VH, and $tt$H production modes.

Higgs boson production in association with single top quark has been studied to derive constraints on the magnitude and relative sign of Higgs boson couplings to top quarks and vector bosons.
Constaints on $\kappa_t$ can be derived using a likelihood ratio scan of $L(\kappa_t)/L(\hat{\kappa_t})$ where $\hat{\kappa_t}$ is the best fit value of $\kappa_t$ as shown in Figure~\ref{figureKt}.
Multilepton final state and final states with single lepton with a pair of bottom quarks were combined with Higgs decay to two photons to get the final result. The data favour positive value of top quark yukawa coupling for SM Higgs boson and excludes ranges of about [-0.9,-0.5] and [1.0,2.1] times $y^{SM}_t$ at the 95 $\%$ confidence level.

\newpage
\section{PROPOSAL}

In this project, it is proposed to search for the production of Higgs boson associated with a single top quark in the CMS detector based at the Large Hadron Collider of the CERN laboratory.
It is proposed to study the sensitivity with the current data in the channel where the events are identified with two muons of the same sign or with three leptons of electron or muon type.
Leptons of tau type are not considered due to complications in their identificaiton in the detector.
%The impact of uncertainties will be studied and projections will be  made to the future phases of the LHC.
For these studies the data of the LHC Run 2 (2015-2018) will be used corresponding to an integrated luminosity of about 150 fb$^{-1}$, this is 4 times larger than the dataset used for current results.
It is also expected to estimate projections of the sensitivity to the High Luminosity phase of the LHC (HL-LHC) (2026-2036) which will accumulate up to 100 times the data used for current measurements.

@@

%- motivation, tH production crossection, sign of $k_t$, search for BSM, ..

There is currently a great enigma in physics since we know that ordinary matter comprises only $5\%$ of the universe, another  $27\%$ includes what we call Dark Matter.
Understanding the production of the Higgs boson, as well as its decays like the invisible decay modes are an important part of the physics program of CERN laboratory experiments to verify the Standard Model and look for new particles.
Through this project we will investigate the production of the Higgs boson in association with a single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC.
The tH  production process has not been observed so far.
The measurement  of this process complements other measurements of the yukawa coupling parameter of the top  quark and Higgs boson and is a part of the physics program of CMS  and ATLAS international collaborations.
Through these measurements we seek to complete the measurements of the Standard Model parameters to high precision or to find evidence of deviations which point to new particles which may explain current enigmas like the origin of Dark Matter.

\section{HYPOTHESIS}
%%  - what we expect to find

Studies in the tH channel have already started with the collaborations CMS and ATLAS with publications based on data from the 2011-2012 run and some data from the second run of 2015-2016.
In these results it was not found evidence of signal and upper limits have been set on the production cross-section.
The current results combine three types of searches: H$\rightarrow WW/ZZ/\tau\tau \rightarrow$leptons, H$\rightarrow\gamma\gamma$, and H$\rightarrow b\bar{b}$.
%The most sensitive channel is  with leptons, however this channel is a combination of two sub-channels: two-lepton and three-lepton.
The currently published result uses 36 fb$^{-1}$ of collision data and obtains an upper limit of 2.04 pb while only 0.077 pb are expected in the SM \cite{ref:tH2019}, the limit is 26 times larger than the expected value.
In BSM models, for example with $\kappa_t = -1$,  the expected cross-section increases up to 0.834 pb due to quantum interference effects; in this case also the observed upper limit changes to 0.86 pb at 95$\%$ confidence level, this result almost rules out this model.

With this project we propose to incorporate the full Run 2 dataset corresponding to 150 pb$^{-1}$.
A factor of 4 increase in number of collision events will improve the limits by a factor of approximately 2 due to the nature of statistical uncertainties.
These means the limit for the SM case will improve to about 13 times the expected cross-section, while for the modified model ($\kappa_t = -1$) we expect to be almost fully excluded.

Additionally, in subsequent phases of the LHC (2021-2023 and 2026-2038) the amount of data will increase by a factor of up to 100, which implies the upper limit for the SM case will reduce to about 2.6 times the expected cross-section. This implies that even for the HL-LHC it will be difficult to observe the tH production, but at the same time more and more phase space of BSM models will be excluded.

In this search the reconstruction methods do not allow to separate $t$H from $t\bar{t}$H production because the cross-section of $t\bar{t}$H is much larger than that of tH and some events pass the selections designed for tH.
The fraction of tH events over tH + $t\bar{t}$H is currently only 5$\%$ with the analysis techniques used.
We expect to improve the analysis techniques using multi-variate techniques, for example boosted decision trees or other algorithms, to improve the identification of signal events over the rest of the backgrounds.
This will provide additional improvements in the limits and a possible observation of the SM signal as well as stringent exclusions for BSM models.

@@

Some of the objectives in this project include:
\begin{itemize}
\item We will study the distribution of events in one of the decay channels and the  sensitivity that we get with the current data.
\item study one or more decay channels.
\item verify that simulation of signal events is adequate.
\item study the extraction of signal strength through adjustments to the data.
\item study the impact of statistical and systematic uncertainties.
\end{itemize}
